
% Terms chapter -------------------------------------------------------
\chapter*{Terms}
\addcontentsline{toc}{chapter}{Terms}
\label{ch:Terms}
This section defines some terms as used in this book.

\section*{Turn}
\addcontentsline{toc}{section}{Turn}
\label{sec:Terms/Turn}
Turn denotes player who hasn't finished his (or her) move, i.e. the one who "has the move",
the one who "is on the move".

\section*{Move}
\addcontentsline{toc}{section}{Move}
\label{sec:Terms/Move}
Move is completed movement of all pieces, performed sequentialy, by one player, in one turn.

\section*{Momentum}
\addcontentsline{toc}{section}{Momentum}
\label{sec:Terms/Momentum}
Momentum is count of step-fields traveled over by a piece.

\section*{Cascading move}
\addcontentsline{toc}{section}{Cascading move}
\label{sec:Terms/Cascading move}
Cascading move is move where at least 2 pieces are moved.

\section*{Ply}
\addcontentsline{toc}{section}{Ply}
\label{sec:Terms/Ply}
Ply is completed movement of a piece, from its' starting position to its' destination field.

\section*{Step-fields}
\addcontentsline{toc}{section}{Step-fields}
\label{sec:Terms/Step-fields}
Step-fields are all fields where a piece can end it's movement.

\section*{Step}
\addcontentsline{toc}{section}{Step}
\label{sec:Terms/Step}
Step is a movement of a piece from one step-field to another.

\section*{Rush}
\addcontentsline{toc}{section}{Rush}
\label{sec:Terms/Rush}
Rush is Pawn's longer initial movement, i.e. from its’ starting position, for at least 2 fields forward.
Rushing Pawn presents opponent with en passant opportunity.

\section*{Capture-fields}
\addcontentsline{toc}{section}{Capture-fields}
\label{sec:Terms/Capture-fields}
Capture-fields are all fields where a piece can capture opponent's piece.
Usually, these are the same as step-fields, except for Pawn.

% \section*{Royal-fields}
% \addcontentsline{toc}{section}{Royal-fields}
% \label{sec:Terms/Royal-fields}
% Royal-fields are all fields surrounding a particular field horizontally,
% vertically and diagonally, i.e. they are step-fields of a King.

\section*{Portal-fields}
\addcontentsline{toc}{section}{Portal-fields}
\label{sec:Terms/Portal-fields}
Portal-fields are all fields immediately surrounding a particular field horizontally,
vertically and diagonally.

\section*{Neighboring-fields}
\addcontentsline{toc}{section}{Neighboring-fields}
\label{sec:Terms/Neighboring-fields}
Neighboring-fields are all fields immediately surrounding a particular field horizontally
and vertically.

\section*{Chip}
\addcontentsline{toc}{section}{Chip}
\label{sec:Terms/Chip}
Chip is a device not playable by either player, used to denote various states
of a game. For instance, it could be used to denote King's ability to castle,
which Pawn is to be promoted, or which player's turn is ongoing.

\section*{Piece}
\addcontentsline{toc}{section}{Piece}
\label{sec:Terms/Piece}
Piece is an item on chessboard, playable by either a player or both players,
or Star, i.e. piece is any item except chips.

\section*{Figure}
\addcontentsline{toc}{section}{Figure}
\label{sec:Terms/Figure}
Figure is any piece, except Pawn.

\section*{Pawn row}
\addcontentsline{toc}{section}{Pawn row}
\label{sec:Terms/Pawn row}
Pawn row is row that contains Pawns on initial setup of chessboard.
For light player that is second row, for dark player second to last row.

\section*{Figure row}
\addcontentsline{toc}{section}{Figure row}
\label{sec:Terms/Figure row}
Figure row is row that contains figures on initial setup of chessboard.
For light player that is first row, for dark player it is last row.

\section*{Piece row}
\addcontentsline{toc}{section}{Piece row}
\label{sec:Terms/Piece row}
Piece row is either Pawn row or figure row.

\section*{Rush row}
\addcontentsline{toc}{section}{Rush row}
\label{sec:Terms/Rush row}
Rush row is any row from which Pawn can be rushed.
Currently, rush rows are the same as piece rows.

\clearpage % ..........................................................
% ------------------------------------------------------- Terms chapter
